An Open-Source Implementation of the CMPS Algorithm for Assessing Similarity of Bullets

### 2022-09-29 rj_format/rj-format-ju-hofmann/ju-hofmann.tex
updated with mathjax

\abstract{%
In this paper, we introduce the R package \pkg{cmpsR}, an open-source
implementation of the Congruent Matching Profile Segments (CMPS) method
developed at the National Institute of Standards and Technology (NIST)
for objective comparison of striated tool marks. The functionality of
the package is showcased by examples of bullet signatures that come with
the package. Graphing tools are implemented in the package as well for
users to assess and understand the CMPS results. Initial tests were
performed on bullet signatures generated from two sets of 3D scans in
the Hamby study under the framework suggested by the R package
\pkg{bulletxtrctr}. New metrics based on CMPS scores are introduced and
compared with existing metrics. A measure called Sum of squares ratio is
included, and how it can be used for evaluating different scans,
metrics, or parameters is showcased with the Hamby study data sets. An
open-source implementation of the CMPS algorithm makes the algorithm
more accessible, generates reproducible results, and facilitates further
studies of the algorithm such as methods comparison.
}

@@

In this paper, we present an open-source implementation of the algorithm
of the Congruent Matching Profile Segments (CMPS) method. \citet{cmps}
developed the CMPS method for ``objective comparison of striated tool
marks'' and demonstrated its use in some examples of comparing bullet
signature correlations. Although \citet{cmps} conceptually describe the
CMPS algorithm in their paper, the authors did not release an
implementation of their method. Thus, our effort here is to introduce
the CMPS method to the \texttt{R} \citep{R} community and provide an
open-source, publicly available implementation of the algorithm to use,
review, and improve. Our implementation is made available as part of the
\texttt{R} package \CRANpkg{cmpsR} on CRAN.

According to the Uniform Crime Reporting Program of the FBI (Federal
Bureau of Investigation), ``more than 76 percent (76.7) of the homicides
for which the FBI received weapons data in 2020 involved the use of
firearms'' \citep{fbiucr}. At the same time, the number of murder
victims jumped between 2019 and 2020 by more than 23 percent (23.4) to
17754 \citep{fbiucr20}. This increase is unprecedented and highlights
the important role that firearm examination plays in all these cases. An
important task of firearm examination is to answer the question of
whether two pieces of evidence come from the same source or whether a
piece of evidence matches a sample obtained from a specific firearm.
Here, in particular, we are interested to determine whether two bullets
were fired from the same gun barrel. Assessing the similarity between
two bullets is based on a comparison of striation marks acquired during
the firing process as bullets are propelled through the barrel. The
current state of art sees firearms examiners make an assessment of
similarity based on a visual comparison, generally, using a comparison
microscope \citep{afte}. This practice has been criticized for its lack
of objectivity and the associated problem of determining valid error
rates \citep{pcast}.

@@

The automated framework for extracting signatures from x3p files used in
this paper was proposed by \citet{aoas} and is implemented in the
\texttt{R} packages \CRANpkg{x3ptools} \citep{x3ptools} and
\pkg{bulletxtrctr} \citep{bulletxtrctr}. \pkg{x3ptools} is a package to
read, write, and generally, process x3p files. The \pkg{bulletxtrctr}
package implements a pipeline for extracting and comparing signatures
from scans of land-engraved areas.

Figure \ref{fig:process} gives an overview of all of the steps in the
process from scan to signatures. Figure \ref{fig:process}(a) shows a
rendering of a 3D scan of a bullet land engraved area. The raised
portion of the surface on the left and right of the scan are parts of
the adjacent groove engraved areas (GEAs), the middle area shows a
land-engraved area (LEA) with well expressed striation marks. The first
step of obtaining the bullet signature is to extract a cross-section at
a fixed height on the land engraving.

The thin white horizontal line in Figure \ref{fig:process}(a) indicates
which cross-section was identified by the algorithm to represent the
LEA; Figure \ref{fig:process}(b) shows the corresponding cross-sectional
view. Groove engraved areas are removed from the analysis as indicated
by the vertical blue lines in Figure \ref{fig:process}(c) and Figure
\ref{fig:process}(d)). A non-parametric LOESS smooth \citep{loess} is
fitted to capture the bullet curvature ( Figure \ref{fig:process}(e))
and, finally, the \textbf{bullet signature} (Figure
\ref{fig:process}(f)) is obtained as residuals of the cross-section and
the fitted smooth. Note, that in \citet{cmps} bullet signatures are
referred to as bullet profiles. However, to avoid confusion, we
distinguish the notion of bullet signatures from bullet profiles. Bullet
profiles are shown in panels (b), (c), and (d) of Figure
\ref{fig:process}, while Figure \ref{fig:process}(f) shows the
corresponding bullet signature. Identifying the groove engraved areas
correctly is fundamental for a correct downstream analysis of the
signatures. To allow for a human-in-the-middle inspection and
intervention we have provided an interactive web application,
implemented as \texttt{R} Shiny App \citep{shiny}, \pkg{bulletinspectR}
to identify and correct those errors. An example of the extraction
process with corresponding code and parameter settings can be found in
the ``Supplementary materials''. Note that the process of extracting
signatures might be different from the one used in \citet{cmps} because
no code or parameter settings are made available publicly.

@@

Most algorithms for comparing striation marks are based on the digitized
signatures and produce a similarity score
\citep[\citet{ChumbleyL_Scott2010VoTM}, \citet{aoas},
\citet{pmid30444940}]{song2005}. The congruent matching profile segments
(CMPS) algorithm, developed by \citet{cmps} for ``objective comparison
of striated tool marks'', is one such algorithm. The algorithm's main
idea is to take a set of consecutive and non-overlapping basis segments
from the comparison and for each segment find the ``best'' registration
position on the reference (the other bullet signature) with respect to
their cross-correlation values. From a comparison of these registration
positions, a \textbf{congruent registration position} is identified, and
the number of basis segments taking the congruent registration position
is the CMPS score. Note that researchers in \citet{cmps} refer to the
origin of basis segments as the reference, but in this paper we refer to
it as the comparison signature. High CMPS scores are achieved between
more similar signatures and are therefore indicative of a same-source
pair. Low scores between pairs of signatures are attributed to different
source pairs. However, a specific threshold of the CMPS score to
distinguish between same-source and different-source comparisons is not
provided in \citet{cmps}, instead the threshold depends on the
underlying structure of the data and the choice of parameters. In a
legal setting this variability is problematic because it allows for
situations in which experts could choose parameters based on whether
they are witnesses for the defense or the prosecution. Further research
is needed to understand how to determine optimal threshold settings. The
CMPS algorithm can assist firearm examiners with drawing a conclusion
about the source of a comparison pair. Thus, in this paper we present an
open-source implementation of the CMPS algorithm in the R package
\pkg{cmpsR} available from both CRAN and Github. This publicly available
implementation calculates the CMPS score of a comparison using the
following code:

@@

The section ``Implementation'' will go through the algorithm and show
how to use the \pkg{cmpsR} package. A further example that illustrates
the main points is also included. In the section on 'Evaluation
metrics'', we introduce new CMPS metrics that summarize land-level CMPS
scores and a sum of squares ratio that can be used to evaluate
algorithmic results. The section ``Results'' presents the results of
evaluating the \pkg{cmpsR} package using Hamby set 252 and Hamby set 44.
The results from Hamby set 252 are used to verify that our
implementation is, at least qualitatively, comparable to the algorithm
described in \citet{cmps}. Results from Hamby 44 show the need for a
further investigation of the parameter choices even in the case of
bullets fired from the same barrels. The last section covers some final
discussion and conclusions.

@@

Note that there are several parameters in the CMPS algorithm that will
affect the final results and are left to the users to decide, such as
the length of basis segments \texttt{seg\_length} in step 1, the number
of peaks \texttt{npeaks\_set} identified on each level in step 2, and
the length of the tolerance zone \texttt{Tx} in both step 2 and 3. In
our implementation of the CMPS algorithm, we used the parameters given
in the original CMPS paper \citep{cmps} as the default values for these
parameters. However, the authors state that no cross-validation has been
done - there might also be issues with respect to the resolution of the
scans. Further research is needed, until then users are advised to think
of default values as starting values and consider alternatives. However,
the evaluation framework based on the Sum of Squares Ratio introduced in
the later section could be used to evaluate the choices of these
parameters. The main function that combines all steps in the CMPS
algorithm described above is called
\texttt{extract\_feature\_cmps(...)}. Here we present it with its
default parameters.

@@

\caption[Signatures of all lands of bullet 1 in the top row, and of bullet 2 in the bottom row]{Signatures of all lands of bullet 1 in the top row, and of bullet 2 in the bottom row. Signatures in the second row are ordered to be in phase with the signatures above, i.e. matching signatures are displayed on top of each other.}\label{fig:sigs}
\end{figure}
\end{Schunk}

@@

In the first example, \texttt{npeaks\_set} is a vector of three
integers, i.e.~the algorithm uses the multi-segment lengths strategy to
create the result object \texttt{cmps\_with\_multi\_scale}. For
\texttt{cmps\_without\_multi\_scale} each basis segment is linked to the
top 5 candidate positions. We use \texttt{include\ =\ "full\_result"} to
capture all results. In this example, the CMPS is 9 when using multiple
segments, and 12 when using a single segment. As discussed in
\citet{cmps}, using multi-segment lengths strategy can reduce the number
of false positives when identifying candidate positions; however, any
score-based method is walking the line between false positives and false
negatives. As the number of false positives is reduced the number of
false negatives might rise. More discussion and comparisons between the
two versions of the CMPS algorithm will be presented in later sections.
Note, that the multi-segment lengths method is slower because the
algorithm is run once for each segment length.

@@

\begin{figure}[hbt]
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot}The black line shows the comparison signature; each red line segment shows one congruent matching profile segment.}
\vspace{1em}
\begin{adjustbox}{valign=t}
\begin{minipage}{.39\textwidth}

@@

\begin{itemize}
\item
  Figure \ref{fig:segplot} is the \texttt{segment\_plot} for basis
  segment 6 at level one (in its original length). We used
  \texttt{npeaks\_set\ =\ c(5,\ 3,\ 1)} in
  \texttt{extract\_feature\_cmps()} when calculating the CMPS score.
  Therefore the top five peaks are identified in the cross-correlation
  curve at level one. Segment 6 is plotted at the positions where these
  5 peaks are identified with dashed lines in the
  \texttt{segment\_plot}. The solid thick black line shows the segment
  at its original position (which in this example is very close to the
  actual registration position).
\item
  Figure \ref{fig:segplot2} is the \texttt{scale\_ccf\_plot} of basis
  segment 6 at level one. It shows the cross-correlation curve computed
  by the reference signature and the level-one basis segment 6. The five
  highest peaks are marked by dots on the curve.
\end{itemize}

@@

We can continue with the example used in previous sections.
\texttt{bullets} contains bullet signatures of two bullets,
\texttt{bullet1} and \texttt{bullet2}. As mentioned before, each bullet
has 6 land engravings, resulting in 6 bullet signatures. Thus, there are
36 pairwise bullet signature comparisons, resulting in 36 \(c_{ij}\)
values in total. We use multi-segment lengths strategy with default
parameters to compute these CMPS scores, and the result is shown in
Figure \ref{fig:tiles}. We can see that in this example,

@@

The sum of squares ratios are 20.64 and 28.96 for
\(\mathrm{CMPS_{max}}\) and \(\mathrm{\overline{CMPS}_{max}}\),
respectively. This indicates, that even though scores from
\(\mathrm{CMPS_{max}}\) for known-match comparisons are larger than
scores from the averaged version of \(\mathrm{\overline{CMPS}_{max}}\),
these scores achieve a better separation between the two groups of
comparisons.

@@

As seen before, the results suggest that the \(\mathrm{CMPS_{max}}\)
metric produces the least amount of separation compared with the other
three CMPS metrics. They also suggest that the two newly proposed
metrics, \(\mathrm{\overline{CMPS}_{diff}}\) and
\(\mathrm{\overline{CMPS^*}_{diff}}\), lead to equally good or even
better results as \(\mathrm{\overline{CMPS}_{max}}\). Because
\(\mathrm{\overline{CMPS^*}_{diff}}\) relies on a scaled version of CMPS
scores it is more comparable to other similarity scores and summarizes
the out-of-phase background CMPS scores, making it superior to the other
CMPS metrics.

@@

In Figure \ref{fig:param-npeak-plot}, we also included the sum of
squares ratios computed from the random forest scores \citep{aoas} for
different metrics. The random forest model presented in \citet{aoas} was
trained at the Center for Statistics and Applications in Forensic
Evidence (CSAFE) and is publicly available in the R package
\pkg{bulletxtrctr} \citep{bulletxtrctr}. Similar to the CMPS algorithm,
this random forest model produces a score to quantify the similarity of
a land-by-land comparison. The RF scores lie in an interval of {[}0,1{]}
making them most comparable to the scaled CMPS scores. We applied the
logic of \(\mathrm{{CMPS}_{max}}\), \(\mathrm{\overline{CMPS}_{max}}\),
and \(\mathrm{\overline{CMPS}_{diff}}\) to the random forest scores and
obtained results of the random forest model for different metrics. As
shown in Figure \ref{fig:param-npeak-plot}, the random forest model does
not achieve the sum of squares ratio as high as those achieved by the
top CMPS algorithms for Hamby 252, but it does achieve better results
for Hamby 44. This might suggest that inclusion of the CMPS score as an
additional feature in the random forest model training might be
beneficial for an overall separation to determine the source.

@@

\begin{itemize}
\tightlist
\item
  \texttt{hamby*\_result\_generator.R}: these R scripts take the
  processed data of Hamby set 252 and Hamby set 44, generate preliminary
  results used in the paper, and save the results in the \texttt{.csv}
  format in the folder \texttt{data-csv}. Please make sure that the
  package versions of \pkg{bulletxtrctr} and \texttt{cmpsR} meet the
  requirements.
\item
  \texttt{rds\_generator.R}: this R script takes the generated
  \texttt{.csv} files and produce \texttt{CMPSpaper\_results.rds}.
  \texttt{CMPSpaper\_results.rds} is identical to
  \texttt{data/CMPSpaper\_results.rds} and is used generate figures and
  other results presented in the paper.
\end{itemize}

>>>>

\abstract{%
In this paper, we introduce the R package \pkg{cmpsR}, an open-source
implementation of the Congruent Matching Profile Segments (CMPS) method
developed at the National Institute of Standards and Technology (NIST)
for objective comparison of striated tool marks. The functionality of
the package is showcased by examples of bullet signatures that come with
the package. Graphing tools are implemented in the package as well for
users to assess and understand the CMPS results. Initial tests were
performed on bullet signatures generated from two sets of 3D scans in
the Hamby study under the framework suggested by the R package
\texttt{bulletxtrctr}. New metrics based on CMPS scores are introduced
and compared with existing metrics. A measure called sum of squares
ratio is included, and how it can be used for evaluating different
scans, metrics, or parameters is showcased with the Hamby study data
sets. An open-source implementation of the CMPS algorithm makes the
algorithm more accessible, generates reproducible results, and
facilitates further studies of the algorithm such as method comparisons.
}

@@

In this paper, we present an open-source implementation of the algorithm
of the Congruent Matching Profile Segments (CMPS) method. \citet{cmps}
developed the CMPS method for ``objective comparison of striated tool
marks'' and demonstrated its use in some examples of comparing bullet
signature correlations. Although \citet{cmps} conceptually describe the
CMPS algorithm in their paper, the authors did not release an
implementation of their method. Thus, our effort here is to introduce
the CMPS method to the \texttt{R} community and provide an open-source,
publicly available implementation of the algorithm to use, review, and
improve. Our implementation is made available as part of the \texttt{R}
package \CRANpkg{cmpsR} on CRAN.

According to the Uniform Crime Reporting Program of the FBI (Federal
Bureau of Investigation), ``more than 76 percent (76.7) of the homicides
for which the FBI received weapons data in 2020 involved the use of
firearms'' \citep{fbiucr}. At the same time, the number of murder
victims jumped between 2019 and 2020 by more than 23 percent (23.4) to
17754 \citep{fbiucr20}. This increase is unprecedented and highlights
the important role that firearm examination plays in all these cases. An
important task of firearm examination is to answer the question of
whether two pieces of evidence come from the same source or whether a
piece of evidence matches a sample obtained from a specific firearm.
Here, in particular, we are interested to determine whether two bullets
were fired from the same gun barrel. Assessing the similarity between
two bullets is based on a comparison of striation marks acquired during
the firing process as bullets are propelled through the barrel. The
current state of the art sees firearms examiners make an assessment of
similarity based on a visual comparison, generally, using a comparison
microscope \citep{afte}. This practice has been criticized for its lack
of objectivity and the associated problem of determining valid error
rates \citep{pcast}.

@@

The automated framework for extracting signatures from x3p files used in
this paper was proposed by \citet{aoas} and is implemented in the
\texttt{R} packages \CRANpkg{x3ptools} \citep{x3ptools} and
\texttt{bulletxtrctr} \citep{bulletxtrctr}. \pkg{x3ptools} is a package
to read, write, and generally, process x3p files. The
\texttt{bulletxtrctr} package implements a pipeline for extracting and
comparing signatures from scans of land-engraved areas.

Figure \ref{fig:process} gives an overview of all of the steps in the
process from scan to signatures. Figure \ref{fig:process}(a) shows a
rendering of a 3D scan of a bullet land engraved area. The raised
portion of the surface on the left and right of the scan are parts of
the adjacent groove engraved areas (GEAs), the middle area shows a land
engraved area with well expressed striation marks. The first step of
obtaining the bullet signature is to extract a cross-section at a fixed
height on the land engraving.

The thin white horizontal line in Figure \ref{fig:process}(a) indicates
which cross-section was identified by the algorithm to represent the
LEA; Figure \ref{fig:process}(b) shows the corresponding cross-sectional
view. Groove engraved areas are removed from the analysis as indicated
by the vertical blue lines in Figure \ref{fig:process}(c) and Figure
\ref{fig:process}(d). A non-parametric LOESS smooth \citep{loess} is
fitted to capture the bullet curvature (Figure \ref{fig:process}(e))
and, finally, the \textbf{bullet signature} (Figure
\ref{fig:process}(f)) is obtained as residuals of the cross-section and
the fitted smooth. Note that in \citet{cmps} bullet signatures are
referred to as bullet profiles. However, to avoid confusion, we
distinguish the notion of bullet signatures from bullet profiles. Bullet
profiles are shown in panels (b), (c), and (d) of Figure
\ref{fig:process}, while Figure \ref{fig:process}(f) shows the
corresponding bullet signature. Identifying the groove engraved areas
correctly is fundamental for a correct downstream analysis of the
signatures. We have provided an interactive web application to allow for
a human-in-the-middle inspection and intervention. It is implemented as
an \texttt{R} Shiny App \citep{shiny} named \texttt{bulletinspectR} for
identifying and correcting those errors. An example of the extraction
process with corresponding code and parameter settings can be found in
the ``Supplementary materials''. Note that the process of extracting
signatures might be different from the one used in \citet{cmps} because
no code or parameter settings are made available publicly.

@@

Most algorithms for comparing striation marks are based on the digitized
signatures and produce a similarity score
\citep[\citet{ChumbleyL_Scott2010VoTM}, \citet{aoas},
\citet{pmid30444940}]{song2005}. The congruent matching profile segments
(CMPS) algorithm, developed by \citet{cmps} for ``objective comparison
of striated tool marks'', is one such algorithm. The algorithm's main
idea is to take a set of consecutive and non-overlapping basis segments
from the comparison and for each segment find the ``best'' registration
position on the reference (the other bullet signature) with respect to
their cross-correlation values. From a comparison of these registration
positions, a \textbf{congruent registration position} is identified, and
the number of basis segments taking the congruent registration position
is the CMPS score. Note that researchers in \citet{cmps} refer to the
origin of basis segments as the reference, but in this paper we refer to
it as the comparison signature. High CMPS scores are achieved between
more similar signatures and are therefore indicative of a same-source
pair. Low scores between pairs of signatures are attributed to different
source pairs. However, a specific threshold of the CMPS score to
distinguish between same-source and different-source comparisons is not
provided in \citet{cmps}, instead the threshold depends on the
underlying structure of the data and the choice of parameters. In a
legal setting this variability is problematic because it allows for
situations in which experts could choose parameters based on whether
they are witnesses for the defense or the prosecution. Further research
is needed to understand how to determine optimal threshold settings. The
CMPS algorithm can assist firearms examiners with drawing a conclusion
about the source of a comparison pair. Thus, in this paper we present an
open-source implementation of the CMPS algorithm in the R package
\pkg{cmpsR} available from both CRAN and Github. This publicly available
implementation calculates the CMPS score of a comparison using the
following code:

@@

The section ``Implementation'' will go through the algorithm and show
how to use the \pkg{cmpsR} package. A further example that illustrates
the main points is also included. In the section on ``Evaluation
metrics'', we introduce new CMPS metrics that summarize land-level CMPS
scores and a sum of squares ratio that can be used to evaluate
algorithmic results. The section ``Results'' presents the results of
evaluating the \pkg{cmpsR} package using Hamby set 252 and Hamby set 44.
The results from Hamby set 252 are used to verify that our
implementation is, at least qualitatively, comparable to the algorithm
described in \citet{cmps}. Results from Hamby 44 show the need for a
further investigation of the parameter choices even in the case of
bullets fired from the same barrels. The last section covers some final
discussion and conclusions.

@@

Note that there are several parameters in the CMPS algorithm that will
affect the final results and are left to the users to decide, such as
the length of basis segments \texttt{seg\_length} in step 1, the number
of peaks \texttt{npeaks\_set} identified on each level in step 2, and
the length of the tolerance zone \texttt{Tx} in both step 2 and 3. In
our implementation of the CMPS algorithm, we used the parameters given
in the original CMPS paper \citep{cmps} as the default values for these
parameters. However, the authors state that no cross-validation has been
done - there might also be issues with respect to the resolution of the
scans. Further research is needed, until then users are advised to think
of default values as starting values and consider alternatives. However,
the evaluation framework based on the sum of squares ratio introduced in
the later section could be used to evaluate the choices of these
parameters. The main function that combines all steps in the CMPS
algorithm described above is called
\texttt{extract\_feature\_cmps(...)}. Here we present it with its
default parameters.

@@

\caption[Signatures of all lands of bullet 1 in the top row, and of bullet 2 in the bottom row]{Signatures of all lands of bullet 1 in the top row, and of bullet 2 in the bottom row. Signatures in the second row are ordered to be in phase with the signatures above, i.e. matching signatures are displayed on top of each other. On the x-axis is the length of the scan in millimeter, and on the y-axis is the relative height in micron.}\label{fig:sigs}
\end{figure}
\end{Schunk}

@@

In the first example, \texttt{npeaks\_set} is a vector of three
integers, i.e.~the algorithm uses the multi-segment lengths strategy to
create the result object \texttt{cmps\_with\_multi\_scale}. For
\texttt{cmps\_without\_multi\_scale} each basis segment is linked to the
top 5 candidate positions. We use \texttt{include\ =\ "full\_result"} to
capture all results. In this example, the CMPS is 9 when using multiple
segments, and 12 when using a single segment. As discussed in
\citet{cmps}, using multi-segment lengths strategy can reduce the number
of false positives when identifying candidate positions; however, any
score-based method is walking the line between false positives and false
negatives. As the number of false positives is reduced the number of
false negatives might rise. More discussion and comparisons between the
two versions of the CMPS algorithm will be presented in later sections.
Note that the multi-segment lengths method is slower because the
algorithm is run once for each segment length.

@@

\begin{figure}[hbt]
\begin{subfigure}[t]{\textwidth}
\caption{\label{fig:sigplot}The black line shows the comparison signature; each red line segment shows one congruent matching profile segment. Each grey rectangle highlights one congruent matching profile segment.}
\vspace{1em}
\begin{adjustbox}{valign=t}
\begin{minipage}{.39\textwidth}

@@

\begin{itemize}
\item
  Figure \ref{fig:segplot} is the \texttt{segment\_plot} for basis
  segment 6 at level one (in its original length). We used
  \texttt{npeaks\_set\ =\ c(5,\ 3,\ 1)} in
  \texttt{extract\_feature\_cmps()} when calculating the CMPS score.
  Therefore the top five peaks are identified in the cross-correlation
  curve at level one. Segment 6 is plotted at the positions where these
  five peaks are identified with dashed lines in the
  \texttt{segment\_plot}. The solid thick black line shows the segment
  at its original position (which in this example is very close to the
  actual registration position).
\item
  Figure \ref{fig:segplot2} is the \texttt{scale\_ccf\_plot} of basis
  segment 6 at level one. It shows the cross-correlation curve computed
  by the reference signature and the level-one basis segment 6. The five
  highest peaks are marked by dots on the curve.
\end{itemize}

@@

We can continue with the example used in previous sections.
\texttt{bullets} contains bullet signatures of two bullets,
\texttt{bullet1} and \texttt{bullet2}. As mentioned before, each bullet
has six land engravings, resulting in six bullet signatures. Thus, there
are 36 pairwise bullet signature comparisons, resulting in 36 \(c_{ij}\)
values in total. We use multi-segment lengths strategy with default
parameters to compute these CMPS scores, and the result is shown in
Figure \ref{fig:tiles}. We can see that in this example,

@@

The sum of squares ratios are 20.64 and 28.96 for
\(\mathrm{CMPS_{max}}\) and \(\mathrm{\overline{CMPS}_{max}}\),
respectively. This indicates that even though scores from
\(\mathrm{CMPS_{max}}\) for known-match comparisons are larger than
scores from the averaged version of \(\mathrm{\overline{CMPS}_{max}}\),
these scores achieve a better separation between the two groups of
comparisons.

@@

As seen before, the results suggest that the \(\mathrm{CMPS_{max}}\)
metric produces the least amount of separation compared with the other
three CMPS metrics. They also suggest that the two newly proposed
metrics, \(\mathrm{\overline{CMPS}_{diff}}\) and
\(\mathrm{\overline{CMPS^*}_{diff}}\), lead to equally good or even
better results as \(\mathrm{\overline{CMPS}_{max}}\). Because
\(\mathrm{\overline{CMPS^*}_{diff}}\) relies on a scaled version of CMPS
scores, it is more comparable to other similarity scores and summarizes
the out-of-phase background CMPS scores, making it superior to the other
CMPS metrics.

@@

In Figure \ref{fig:param-npeak-plot}, we also included the sum of
squares ratios computed from the random forest scores \citep{aoas} for
different metrics. The random forest model presented in \citet{aoas} was
trained at the Center for Statistics and Applications in Forensic
Evidence (CSAFE) and is publicly available in the R package
\texttt{bulletxtrctr} \citep{bulletxtrctr}. Similar to the CMPS
algorithm, this random forest model produces a score to quantify the
similarity of a land-by-land comparison. The RF scores lie in an
interval of {[}0,1{]} making them most comparable to the scaled CMPS
scores. We applied the logic of \(\mathrm{{CMPS}_{max}}\),
\(\mathrm{\overline{CMPS}_{max}}\), and
\(\mathrm{\overline{CMPS}_{diff}}\) to the random forest scores and
obtained results of the random forest model for different metrics. As
shown in Figure \ref{fig:param-npeak-plot}, the random forest model does
not achieve the sum of squares ratio as high as those achieved by the
top CMPS algorithms for Hamby 252, but it does achieve better results
for Hamby 44. This might suggest that inclusion of the CMPS score as an
additional feature in the random forest model training might be
beneficial for an overall separation to determine the source.

@@

\begin{itemize}
\tightlist
\item
  \texttt{hamby*\_result\_generator.R}: these R scripts take the
  processed data of Hamby set 252 and Hamby set 44, generate preliminary
  results used in the paper, and save the results in the \texttt{.csv}
  format in the folder \texttt{data-csv}. Please make sure that the
  package versions of \texttt{bulletxtrctr} and \texttt{cmpsR} meet the
  requirements.
\item
  \texttt{rds\_generator.R}: this R script takes the generated
  \texttt{.csv} files and produce \texttt{CMPSpaper\_results.rds}.
  \texttt{CMPSpaper\_results.rds} is identical to
  \texttt{data/CMPSpaper\_results.rds} and is used generate figures and
  other results presented in the paper.
\end{itemize}
